\textbf{On the Common Logical Structure of Classical and Quantum Mechanics}

### 2021-06-21 Articles/2020-QuantumLogic/Source/quantumlogic.tex
Finished review

Now that we have argued that quantum mechanics can be given an algebra that looks more like the one of classical mechanics, let us show that we can (at least formally) give classical mechanics an algebra that looks like the one of quantum logic. Given that quantum mechanics is inherently a probabilistic theory, let us compare it to a classical probabilistic theory. Thus, let us consider the space of all possible probability distributions $\rho(x,p)$ over phase-space. As already stated, we can formalize this as a Hilbert space, where a vector $\psi$ corresponds to $\sqrt{\rho(x,p)}$. This will be a real vector space, and we are only going to consider operators that can be expressed as a multiplicative function $f(x,p)$. Note that $\langle \psi | F \psi \rangle = \int f(x,p) \rho(x,p) dx dp$ as one would expect.\footnote{To be clear, we are not arguing that this construction is the most appropriate or useful. Simply that it can be done.}  We can now take the algebra of subspaces and, as in the quantum case, it will be non-distributive: the disjunction of two subspaces will not be the set union, but the span.

@@

Note that during the discussion there was no mention as to whether $k$ and $j$ where real numbers, complex numbers or quaternions. The argument is formally the same in all cases. Therefore the lack of distributivity has nothing to do with the complexity of the Hilbert space or non-commuting observables. It's a property of the lattice of subspaces of any vector space and, more generally, of the lattice of subgroups of any group. Indeed, as a further example, consider the group of boost transformations in relativity theory. The boosts along one direction form a subgroup. If we take the subgroups of boost along $x$ and $y$ respectively, the smallest subgroup that contain both, the disjunction, is the subgroup of all boosts along any direction in the $x$-$y$ plane. As before, the disjunction is the span, hence the resulting set is different from the union, and the union of all boosts along the $x$-axis and the $y$-axis is not a subgroup and therefore not contained in the lattice. Thus, this has the same non-distributive structure as quantum logic.

@@

More precisely, in this essay we argued that in order to understand the actual meaning of quantum propositions it is important to take into account whether they refer to preparations or measurements. Indeed, it is the temporal ambiguity present in the example discussed in Section \ref{QL} that led to the wrong conclusion according to which the distributive law is refuted by the physical content of standard quantum mechanics. Contrary to this claim, we argued that making explicit the temporal dependence of logical propositions is sufficient to recover \eqref{DL} in the quantum context. In addition, this result entails that the non-commutativity of quantum observables does not play any role in refuting or rejecting the distributivity law, meaning that such peculiar algebraic property of quantum operators do not change the rules of logic, dissolving one of the most important myths of QL. Similarly, we explained that the existence of superpositions does not entail per se that quantum disjunctions are true even in the case in which neither of its members is.

Furthermore, in Section \ref{Math} it has been shown that taking into account the relevant and fundamental algebraic structures on which quantum mechanics is built, i.e.\ the Borel algebra, it is possible to demonstrate that quantum theory can generate a distributive lattice. On the other hand, we also argued that considering the space of probability distributions over phase space one can assign a non-distributive lattice to classical propositions. These facts indicate that the complexity of the Hilbert space characterizing quantum theory, as well as the non-commutativity of quantum observables do not entail any revision of the rules of classical logic in the quantum domain. Referring to this, going back to the very mathematical foundations of both classical and quantum theories, we saw that the Borel algebra, which is a Boolean algebra, characterizes the formal edifice on which both theories are built, and guarantees that both frameworks can be given a distributive lattice of logical propositions, showing as a consequence a remarkable similarity. The non-distributivity of quantum lattices originates from the simple algebraic fact that they do not employ standard set operations. However, this feature is a property of the lattice of subspaces of \emph{any} vector space, and of the lattice of subgroups of any group more generally. Hence, we can safely conclude that the mathematical structures on which quantum theory rest do not entail a rejection of classical logic.

>>>>

Now that we have argued that quantum mechanics can be given an algebra that looks more like [TODO: Should a stronger phrase than "looks more like" be used?] the one of classical mechanics, let us show that we can (at least formally) give classical mechanics an algebra that looks like the one of quantum logic. Given that quantum mechanics is inherently a probabilistic theory, let us compare it to a classical probabilistic theory. Thus, let us consider the space of all possible probability distributions $\rho(x,p)$ over phase-space. As already stated, we can formalize this as a Hilbert space, where a vector $\psi$ corresponds to $\sqrt{\rho(x,p)}$. This will be a real vector space, and we are only going to consider operators that can be expressed as a multiplicative function $f(x,p)$. Note that $\langle \psi | F \psi \rangle = \int f(x,p) \rho(x,p) dx dp$ as one would expect.\footnote{To be clear, we are not arguing that this construction is the most appropriate or useful. Simply that it can be done.}  We can now take the algebra of subspaces and, as in the quantum case, it will be non-distributive: the disjunction of two subspaces will not be the set union, but the span.

@@

Note that during the discussion there was no mention as to whether $k$ and $j$ were real numbers, complex numbers or quaternions. The argument is formally the same in all cases. Therefore the lack of distributivity has nothing to do with the complexity of the Hilbert space or non-commuting observables. It's a property of the lattice of subspaces of any vector space and, more generally, of the lattice of subgroups of any group. Indeed, as a further example, consider the group of boost transformations in relativity theory. The boosts along one direction form a subgroup. If we take the subgroups of boost along $x$ and $y$ respectively, the smallest subgroup that contain both, the disjunction, is the subgroup of all boosts along any direction in the $x$-$y$ plane. As before, the disjunction is the span, hence the resulting set is different from the union, and the union of all boosts along the $x$-axis and the $y$-axis is not a subgroup and therefore not contained in the lattice. Thus, this has the same non-distributive structure as quantum logic.

@@

More precisely, in this essay we argued that in order to understand the actual meaning of quantum propositions it is important to take into account whether they refer to preparations or measurements. Indeed, it is the temporal ambiguity present in the example discussed in Section \ref{QL} that led to the wrong conclusion according to which the distributive law is refuted by the physical content of standard quantum mechanics. Contrary to this claim, we argued that making explicit the temporal dependence of logical propositions is sufficient to recover \eqref{DL} in the quantum context. In addition, this result entails that the non-commutativity of quantum observables does not play any role in refuting or rejecting the distributivity law, meaning that this peculiar algebraic property of quantum operators does not change the rules of logic, dissolving one of the most important myths of QL. Similarly, we explained that the existence of superpositions does not entail per se that quantum disjunctions are true even in the case in which neither of its members is.

Furthermore, in Section \ref{Math} it has been shown that taking into account the relevant and fundamental algebraic structures on which quantum mechanics is built, i.e.\ the Borel algebra, it is possible to demonstrate that quantum theory can generate a distributive lattice. Moreover, we argued that considering the space of probability distributions over phase space one can assign a non-distributive lattice to classical propositions. These facts indicate that neither the complexity of the Hilbert space characterizing quantum theory nor the non-commutativity of quantum observables entails any revision of the rules of classical logic in the quantum domain. Referring to this, going back to the very mathematical foundations of both classical and quantum theories, we saw that the Borel algebra, which is a Boolean algebra, characterizes the formal edifice on which both theories are built, and guarantees that both frameworks can be given a distributive lattice of logical propositions, showing as a consequence a remarkable similarity. The non-distributivity of quantum lattices originates from the simple algebraic fact that they do not employ standard set operations. However, this feature is a property of the lattice of subspaces of \emph{any} vector space, and of the lattice of subgroups of any group more generally. Hence, we can safely conclude that the mathematical structures on which quantum theory rest do not entail a rejection of classical logic.
